### 2016-12-12 paper/paper.tex
added sections to paper

\documentclass[11pt]{article}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{fullpage}
\usepackage{enumerate}
\renewcommand{\theenumi}{\roman{enumi}}
\usepackage{color}
\definecolor{Purple}{rgb}{.9,0,.9}
\newcommand{\solution}[1]{\textcolor{Purple}{\\Solution: #1}}  %Solution
\newcommand{\C}[1]{\ensuremath{\mathord{\rm #1}}}
\newcommand{\pair}[1]{\ensuremath{\mathopen{\langle}#1\mathclose{\rangle}}}
\newcommand{\lng}[1]{\ensuremath{\mathopen{|}#1\mathclose{|}}}
\newcommand{\card}[1]{\ensuremath{\mathopen{|\!|}#1\mathclose{|\!|}}}
\newcommand{\manyone}{\ensuremath{\leq_m^p}}
\newcommand{\pmli}{\ensuremath{\leq_{m,\mathord{\rm li}}^p}}
\newcommand{\ponem}{\ensuremath{\mathrel{\leq_m^{p/1}}}}
\newcommand{\ponett}{\ensuremath{\mathrel{\leq_{1-\mathord{\rm tt}}^{p/1}}}}
\newcommand{\ptt}{\ensuremath{\mathrel{\leq_{\mathord{\rm tt}}^p}}}
\newcommand{\pktt}{\ensuremath{\mathrel{\leq_{k-\mathord{\rm tt}}^p}}}
\newcommand{\pttk}[1]{\ensuremath{\mathrel}{\leq_{#1-\mathord{\rm
        tt}}^p}}
\newcommand{\pmhat}{\ensuremath{\mathrel{\leq_{\hat{m}}^p}}}
\newcommand{\pmhatli}{\ensuremath{\mathrel{\leq_{\hat{m}\mathord{\rm
          ,l.i.}}^p}}}
\newcommand{\pmhathonest}{\ensuremath{\mathrel{\leq_{\hat{m},\mathord{\rm honest}}^p}}}
\newcommand{\PNP}{\C{P}^{\C{NP}}}
\newcommand{\pT}{\ensuremath{\mathrel{\leq_T^p}}}
\newenvironment{proof}{\vspace*{1em}\noindent{\bf Proof.}}{\hfill$\Box$}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\usepackage{graphicx}
\graphicspath{ {./} }
%\setlength{\parindent}{0cm}
\title{Identifying Pathway Modules from High Throughput Genetic Interaction Data using Simulated Annealing}
\author{Jeremy Colebrook-Soucie, Quinn Collins}

\newtheorem{definition}[theorem]{Definition}

% NOTE! Affiliations placed here should be for the institution where the
%       BULK of the research was done. If the author has gone to a new
%       institution, before publication, the (above) affiliation should NOT be changed.
%       The authors 'current' address may be given in the "Author's addresses:" block (below).
%       So for example, Mr. Fogarty, the bulk of the research was done at UIUC, and he is
%       currently affiliated with NASA.

\begin{document}

\maketitle

% Head 1

\section{Abstract}

\section{Introduction}

\par In this paper, we explore simulated annealing as an alternative method for detecting patterns from high throughput genetic interaction data. This work has been done as an extension of the genecentric package and the LocalCut method defined in Leiserson \textit{et al}. We also touch upon subsequent verification techniques.

\par High throughput genetic interaction data can be represented as a weighted, undirected graph, in which each vertex represents a gene and each edge represents how positive or negative the genetic interaction between the two genes is. Thus, we can identify genes that belong in the same pathway because the genes are likely to have positive genetic interactions with one another. We can also identify genes that belong in a related pathway, because the genes in one pathway will have many negative genetic interactions with another related pathway. Thus, we can model finding related genetic pathways as a graphing problem.

\par We represent related genetic pathways as BPMs (between pathway models), which are each composed of two modules of genes. Each module represents a possible gene pathway, and each BPM represents 2 possible related gene pathways. Thus, we aim to maximize the positive edge weights in each module, and negative edge weights between modules in the same BPM.

\par We apply simulated annealing to high throughput genetic interaction data to generate possible BPMs. Applying simulated annealing allows us to generate "bad" BPMs early on, that might not have strong positive interactions within each module and strong negative interactions between modules, in order to explore the sample space. As the process continues, we gradually move towards better BPMs until we find a set of BPMs that have weights within and between modules that we consider "good". Thus, simulated annealing allows us to explore a large possible sample space and come up with a probable solution.

\section{Data}

The data that we examined and performed our analysis on was created by Tufts' Bioinformatics and Computational Biology Research Group. This dataset was based on Collins \textit{et al.}'s genetic interaction map on protein complexes involved in yeast chromosome biology from the Krogan Lab Interactome Database[Collins 2007]. The format of this dataset was consistent with that required for the genecentric package that was used[Leiserson 2011, Gallant 2013]. 
\section{Methods}
\subsection{Simulated Annealing}
\par Simulated annealing is a hueristic that is used for solving optimization problems with a very large sample space. Its function is analogous to the physical process of annealing, where a metal is heated and then slowly cooled in order to remove imperfections. When the metal is hot, the smith can safely make large changes to it. However, as it colder, he must work it more slowly into its final shape.

\par Let us define the following terminology.

\begin{definition}
{\bf Objective Function} The objective function is a function from an element in the sample space to the real numbers. Simulated annealing attempts to minimize this function.
\end{definition}

\begin{definition}
{\bf Permutation} In the context of simulated annealing, a permutation is how one gets from one valid point in the sample space to another. In general, such permutations are simple, random changes in a solution. 
\end{definition}

\begin{definition}
{\bf Temperature} The temperature of simulated annealing defines how willing it is to accept worse solutions. When temperature is high, simulated annealing has a large probability of accepting a solution that is worse than the previous candidate. However, when the temperature is low, simulated annealing will accept worse solutions with lower and lower probabilties. 
\end{definition}

\par Simulated annealing works by exploring the sample space, with the end goal of minimizing the objective function of the final solution. It begins with a very high temperature, which exponentially decays over a fixed number of total iterations. At every iteration, it generates a new solution based on the previous solution. It then  evaluates this new solution and then decides whether or not to accept it based on how it compares to the previous solution and the current temperature.

\par Thus, in order to apply simulated annealing to the problem domain of extracting BPMs from a genetic interaction network, we must define the sample space, and how to both permuate and evaluate a point in the sample space.

\subsection{Sample Space}
\par We defined the sample space as a partition of the nodes in the genetic interation network. Each partition represents a single BPM. Each BPM is in turn divided into two modules.

\par This is a notable divergence from the work of Leiserson \textit{et al}. In this work, genes are allowed to be placed in multiple BPM's at once. We opted not to include this feature because we believed it would have greatly excessively expanded the sample space and would have exacerabated some on the problems inherent to our approach. An unfortunate of this is that it made comparing our work to previous work a little bit harder.

\subsection{Objective Function}
We defined our objective to be the sum of a BPM objective function times a scaling factor across all BPMs.

Define this BPM function as:
\begin{align}
Bpm Obj(x) = \sum_{a \in M_x}^{}  \sum_{b \in M_x}^{} e_{a, b}  
		   + \sum_{a \in N_x}^{}  \sum_{b \in N_x}^{} e_{a, b}  
		   - \sum_{a \in M_x}^{}  \sum_{b \in N_x}^{} e_{a, b}              
\end{align}

$M_x$ and $N_x$ are the two submodules defining BPM x. $e_{a, b}$ is the weight of the edge between a and b. If an edge does not exist, then it is 0.

Define the scaling factor as:
$$scale(i) = \frac{e^{-\frac{(i - \mu) ^2}{2*\sigma^2}}}{\sqrt{2 * \pi} * \sigma}$$
We experimented with multiple $\sigma$ and  $\mu$ values, but eventually settled on $\mu = 30$ and $\sigma = 5$.

Finally, define our objective function as:
$$obj(Bpms) = \sum_{x \in Bpms}{} Bpm Obj(x) * scale(|M_x| + |N_X|) $$

\par We chose the normal distribution as the scaling factor because it is an easily understandable and modifiable bell curve. Such a scaling factor was necessary because all the other objective functions that we tried led to the creation of many modules with a size less than 3 or greater than 25, which were subsequently discarded when pruning results.

\par We completely recognize that using a scaling factor may have caused some problems. We will address these in the discussion section.

\subsection{Permutation Behavior}
\par The procedure for permuting a set of BPMs is as follows.

\par Begin by selecting a random BPM. If that BPM has a module that has a size greater than 25, perform a split. That is, split both modules in have and create two new BPMs created from these two splits. Remove the original BPM.

\par Otherwise, randomly pick one of the two modules and remove a random gene from that module. If removing this gene empties the BPM (as in, both modules are empty), delete the BPM. Next, pick another random BPM and a random module from within it. Add this gene to that module.

\section{Convergence}
To determine how consistent the BPM modules we generated were, we ran our simulated annealing algorithm 100 times, and recorded how often each gene appeared in the same module as each other gene. We then compared these results to a Monte Carlo simulation. To generate the Monte Carlo simulation data, we assigned a number to each gene in our dataset and randomly partitioned the set of gene placeholder numbers into partitions of size 15. We then counted how often each placeholder number appeared in the same module as each other placeholder number. We would expect that if the BPM modules were consistent, then genes would often either appear in the same module as other genes 0 times, or many times. We found, however, that the consistency of our BPMs nearly matched that of the Monte Carlo simulation data. We then recorded the consistency of BPM modules generated by Genecentric using the LocalCut algorithm. We found that Genecentric with LocalCut had a far different distribution, with far more genes that never appear in the same module as other genes, and far more genes that appear often with other genes. The results for Genecentric/LocalCut are not pictured in the graph below because the algorithm allows for each gene to appear multiple times in the final collection of BPMS, which put the consistency results on a different order of magnitude. Thus, because the consistency of BMP modules created by our simulated annealing algorithm closely matched that created by the Monte Carlo simulation, and were distributed differently than that of the Genecentric/LocalCut algorithm, we can conclude that our BPM modules were not consistent.

\includegraphics[scale=0.8]{convergence.png}

\section{Comparison to Genecentric}
%jeremy@mint2 ~/Documents/comp150bg/comp_bio_final_project $ python count_enrichment_dir.py 100_random_enrich/
%0.647129186603 0.0370813397129 28.4736842105, 44
%jeremy@mint2 ~/Documents/comp150bg/comp_bio_final_project $ python count_enrichment_dir.py 100_enrich/
%0.597336658156 0.0313344465786 27.0727272727, 45.6
%jeremy@mint2 ~/Documents/comp150bg/comp_bio_final_project $ python count_enrichments.py %../genecentric/results/yeast_filtered.gobpm 
%(0.9393939393939394, 0.4090909090909091, 62.0, 66)

\par It is not entirely clear how compare genecentric to simulated annealing. This is because genecentric allows for genes to be in mulitple modules, obviously results in more, higher quality modules than simulated annealing. Ultimately, we decided to lean on percent of BPMs enriched and percent of BPMs enriched for same function to standardize these measurements.

\par This line of thought led us to the conclusion that we would need a random group, where BPMs were randomly generated, to compare our results to. Each random group contained about 22 BPMs where each module contained 15 random genes. The average over many random trials would form a baseline. In the end, we took the average over FIXME FIXME FIXME.

\par Additionally, to correct for the apparent lack of convergence in the simulated annealing method, we took the average results over 55 trials.

\par In order to compare these different methods, we examined several different benchmarks. Modules found is the number of modules found between size 3 and 25. Modules enriched is the number of modules enriched for at least one GO term at consider  $p \leq 0.05 $. BPMs Enriched for Same Function is the number of BPMs where both modules had at least one shared GO term.
\begin{center}
\begin{tabular}{ |c | c |c| c | } 
 \hline
 Method & Modules Found & Modules Enriched & BPMs Enriched for Same Function \\
	\hline
 Genecentric              & 66 & 62 (94\%) & 27 (41\%)\\ 
	\hline 
 Simulated Anneal Average & 45.6 & 27.1 (59.7\%) & 1.4 (0.031\%)\\ 
 \hline
 Random Partition Average & 44.0 & 26.47 (64.7\%) & 1.5 (0.037\%)\\ 
 \hline
\end{tabular}
\end{center}
\par This data is particuarly telling. As anticipated, genecentric performed very well. Consequently, our method certainly did not, performing at the same level as completely random selection. This will be explored further in the discussion section.

\section{Discussion}
Postmortem about why SA bad in this domain

\section{References}
Collins S. Miller K. Maas N., et al. Functional dissection of protein complexes involved in yeast chromosome biology using a genetic interaction map. Nature. 2007;446:806–810.\\

M. Leiserson, D. Tatar, L. Cowen and B. Hescott, "Inferring Mechanisms of Compensation from E-MAP and SGA Data Using Local Search Algorithms for Max Cut," Recomb 2011, pp. 154-167.\\

Gallant, Leiserson, Kackalov, Cowen and Hescott, "Genecentric: A package to uncover graph-theoretic structure in high-throughput epistasis data." BMC Bioinformatics in press, 2013.\\

\end{document}
% End of v2-acmlarge-sample.tex (March 2012) - Gerry Murray, ACM